---
title: "R Notebook"
output: html_notebook
---

### Import required data and libraries

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# clear the R environment
rm(list=ls())

library(tidyverse)
library(latex2exp)
library(scales)
library(extrafont)
library(viridis)
library(ggridges)
library(forcats)
library(ggExtra)
library(grid)
library(gtable)
# set fonts
loadfonts(device = "pdf")
# font_import()
# link www.fontsquirrel.com/fonts/latin-modern-roman

# execute once to add fonts:
# font_import(pattern = "lmroman*")
# theme(legend.position = "top", text=element_text(size=14, family="LM Roman 10"))

# set ggplot global theme
theme_set(theme_bw() +
            theme(legend.position = "top") +
            theme(text = element_text(size = 16, family = "LM Roman 10")))

# multiple figures together
if(!require(devtools)) install.packages("devtools")
devtools::install_github("thomasp85/patchwork")
library(patchwork)

# load pre-processed data
load(file="Data/description.RData")
load(file="Data/results.RData")
data <- read_csv("Data/data.csv", col_names = T)

```

### How much of the used dependencies are explicitly declared by the clients?

```{r}
nb_direct_and_transitive_deps <- 
  results %>%
  dplyr::group_by(Artifact, Type) %>%
  dplyr::summarize(n = n()) %>%
   mutate(percentage = n / sum(n)*100) %>% 
  ungroup() %>% 
  tidyr::complete(Artifact, Type, fill = list(n = 0))


# summary statistics of direct and transitive dependencies
direct_deps <- nb_direct_and_transitive_deps %>% filter(Type == "direct")
transitive_deps <- nb_direct_and_transitive_deps  %>% filter(Type == "transitive")
summary(direct_deps$n)
summary(transitive_deps$n)

print("percentage of direct deps:")
(sum(direct_deps$n) * 100) / (sum(direct_deps$n) + sum(transitive_deps$n))
print("percentage of transitive deps:")
(sum(transitive_deps$n) * 100) / (sum(direct_deps$n) + sum(transitive_deps$n))

nb_direct_and_transitive_deps$Type <- ifelse(nb_direct_and_transitive_deps$Type == "direct", "Direct", "Transitive")

# save the data objects
save(nb_direct_and_transitive_deps, file = "Data/nb_direct_and_transitive_deps.RData")
load(file = "Data/nb_direct_and_transitive_deps.RData")

```


```{r}
direct_n <- nb_direct_and_transitive_deps %>% filter(Type == "Direct") %>% dplyr::rename(Direct = n)
transitive_n <- nb_direct_and_transitive_deps %>% filter(Type == "Transitive") %>% dplyr::rename(Transitive = n)

p <- inner_join(direct_n, transitive_n, by = "Artifact") %>% 
  ggplot(aes(x = Direct, y = Transitive)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method='lm',formula=y~x) +
   scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  labs(x = "#Direct dependencies", y = "#Transitive dependencies")

p <- ggExtra::ggMarginal(p, type = "boxplot") 

ggsave(filename = "Figures/nb_direct_and_transitive_deps_boxplot.pdf", plot = p,
       height = 3, width = 7,  units = c("in"), device = "pdf")


```

```{r}
# ----------------------------------------------------------------------------
# Violinplot with the distribution of direct and transitive dependencies
# ----------------------------------------------------------------------------

# plot boxplot the distribution of direct and transitive dependencies
means <- aggregate(n ~  Type, nb_direct_and_transitive_deps, mean)
means$n <- round(means$n, 2)
nb_direct_and_transitive_deps_boxplot <- 
  nb_direct_and_transitive_deps %>% 
  ggplot(aes(Type, n)) +
  geom_violin(trim = FALSE, aes(fill = Type)) +
  geom_boxplot(width = 0.1) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  geom_point(data = means,
             aes(x=Type, y=n),
             colour="black",
                shape=18, size=6, show.legend = FALSE)+
  geom_point(data = means,
             aes(x=Type, y=n),
             colour="lightblue", 
                shape=18, size=4, show.legend = FALSE)+
  geom_text(data = means, aes(label = n, y = n)) +
  scale_fill_viridis(discrete = TRUE, option = "D", alpha = 0.75) +
  ylab(TeX("#Dependencies")) +
  xlab("Type") +
  coord_flip() +
  theme(legend.position = "none")

ggsave(filename = "Figures/nb_direct_and_transitive_deps_boxplot.pdf", plot = nb_direct_and_transitive_deps_boxplot,
       height = 3, width = 7,  units = c("in"), device = "pdf")

# ----------------------------------------------------------------------------
# Geom_area with the distribution of direct and transitive dependencies
# ----------------------------------------------------------------------------

# Calculate the mean of each group
# mu <- ddply(nb_direct_and_transitive_deps, "Type", summarise, grp.mean=mean(percent))
nb_direct_and_transitive_deps$Artifact <- as.numeric(as.factor(nb_direct_and_transitive_deps$Artifact))
nb_direct_and_transitive_deps$Artifact <- as.factor(nb_direct_and_transitive_deps$Artifact)

nb_direct_and_transitive_deps_arranged <- nb_direct_and_transitive_deps %>% 
  group_by(Artifact, Type) %>% 
  arrange(percentage)

nb_direct_and_transitive_deps_area <-
  nb_direct_and_transitive_deps_arranged %>%
  ggplot(aes(
    x = Artifact, 
    y = percentage,
    fill = Type)) +
  geom_area(alpha = 1, position = "fill") +
  ylab("%Dependencies") +
  scale_x_discrete() +
  xlab("Libraries") +
  xlim(0, 10000) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_viridis(discrete = TRUE) 

# deps_area_deps_boxplot <- nb_direct_and_transitive_deps_area + nb_direct_and_transitive_deps_boxplot +plot_layout(ncol = 1)

ggsave(filename = "Figures/nb_direct_and_transitive_deps_area.pdf", plot = nb_direct_and_transitive_deps_area,
       height = 3, width = 7,  units = c("in"), device = "pdf")

```

### How frequent is the occurrence of bloated dependencies in the Maven ecosystem?

Barplot with the Types of Usages

```{r}
results_bloat <- results %>% 
  # filter(Scope == "compile", Optional == "false", Type == "direct") %>% 
  mutate(TypeUsage = ifelse(UsedDeclared == T, "UsedDeclared", ifelse(UsedUndeclared == T, "UsedUndeclared", ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared"))))

# percentage of each scope
percentage_scopes <- results_bloat %>% 
  filter(Scope %in% c("compile", "test", "runtime", "provided")) %>% 
  dplyr::group_by(Scope) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)*100) %>% 
  ungroup()

# percentage of usage types for the compiled scope
filtered_results_bloat<- results_bloat %>% 
  filter(Scope %in% c("compile")) %>% 
  dplyr::group_by(TypeUsage) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)) %>% 
  ungroup() 

filtered_results_bloat$label <- round(filtered_results_bloat$percentage * 100, 1)
 
filtered_type_bloat_barplot <- filtered_results_bloat %>% 
  ggplot(aes(x = "Compile", y = percentage, fill = TypeUsage, label = paste0(label,"%"))) +
    geom_bar(stat="identity") +
    geom_text(size = 5, position = position_stack(vjust = 0.5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(y = "%Dependencies", x = "Scope") +
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip()

ggsave(filename = "Figures/filtered_type_bloat_barplot.pdf", plot = filtered_type_bloat_barplot ,
       height = 3, width = 7,  units = c("in"), device = "pdf")


results_bloat %>% group_by(Artifact, TypeUsage) %>% summarise(n = n()) %>% arrange(desc(n)) %>% 
  filter(Artifact == "org.wso2.carbon:org.wso2.carbon.core:4.4.27")


results_bloat %>% group_by(Artifact, TypeUsage) %>% summarise(n = n()) %>% arrange(desc(n)) %>% 
  filter(Artifact == "org.eclipse.che.core:che-core-api-core:6.10.0")


```

#### Impact of bloat per community

```{r}
description %>% count(Organization) %>% arrange(desc(n))

# filter Apache Software Foundation projects
description %>% filter(str_detect(Artifact, "org.apache"))

# filter Eclipse Foundation artifacts
description %>% filter(str_detect(Artifact, "org.eclipse"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "org.wso2"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.facebook"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.amazon"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.netflix"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.google"))

communities <- results_bloat %>% select(Artifact, AllDeps, TypeUsage) %>% mutate(Community =
                                                           ifelse(
                                                             startsWith(Artifact, "org.apache"),
                                                             "Apache (1103)",
                                                              ifelse(
                                                               startsWith(Artifact, "org.eclipse"),
                                                               "Eclipse (406)", 
                                                               ifelse(startsWith(Artifact, "org.wso2"),
                                                                      "OW2 (536)", 
                                                                      ifelse(
                                                                        startsWith(Artifact, "com.facebook"),
                                                                                   "Facebook (17)", 
                                                                                   ifelse(
                                                                                     startsWith(Artifact, "com.amazon"),
                                                                                                "Amazon (91)",
                                                                                                ifelse(
                                                                                                  startsWith(Artifact, "com.netflix"),
                                                                                                             "Netflix (51)", 
                                                                                                             ifelse(
                                                                                                               startsWith(Artifact, "com.google"), "Google (62)", "Other")
                                                                                                             ))
                                                                                                ))
                                                                                   ))
                                                                      )
# percentage of usage types for per community
communities <- communities %>% 
   filter(Community != "OW2 (536)" & Community != "Other") %>% 
  dplyr::group_by(Community, TypeUsage) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)) %>% 
  ungroup() 

communities$label <- round(communities$percentage * 100, 1)

# communities$TypeUsage <- factor(communities$TypeUsage, levels = rev(levels(communities$TypeUsage)))

communities$Community <- factor(communities$Community,levels = c("Google (62)", "Netflix (51)", "Amazon (91)", "Facebook (17)", "Eclipse (406)", "Apache (1103)" ))

communities_bloat <- communities %>% 
  ggplot(aes(x = Community, y = percentage, fill = TypeUsage, label = paste0(label,"%"))) +
    geom_bar(stat="identity") +
    geom_text(size = 4, position = position_stack(vjust = 0.5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(y = "%Dependencies", x = "Communities") +
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip() 

ggsave(filename = "Figures/communities_bloat.pdf", plot = communities_bloat ,
       height = 3, width = 8,  units = c("in"), device = "pdf")
                                                          
```

#### Distribution of bloated dependencies across dependencies

```{r}
deps_count_types <- results %>%
  dplyr::filter(Scope == "compile" & Optional == F) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>%
  group_by(Artifact, TypeUsage) %>%
  summarise(count_type_usage = n())

total_deps <- results %>% group_by(Artifact) %>% summarize(total_deps = n())

dist_deps_count_bloat_types <- inner_join(deps_count_types, total_deps, by = "Artifact") %>% 
  filter(TypeUsage== "BloatedDeclared" | TypeUsage== "BloatedUndeclared") %>%
  ggplot(aes(x = count_type_usage, y=total_deps, color = TypeUsage, shape = TypeUsage)) +
  geom_point(alpha = 0.5, size = 3) +
  geom_smooth(se=F) +
  # scale_y_log10(
  #   breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
  #   labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  # ) +
  scale_y_continuous(limits=c(0, 5000)) +
  xlab("#Bloated dependencies") +
  ylab("#Dependencies") +
  # scale_color_viridis(discrete = TRUE) +
  scale_color_manual(values = c("#482173", "#2e6f8e")) +
  theme(legend.title = element_blank())

ggsave(filename = "Figures/dist_deps_count_bloat_types.pdf", plot = dist_deps_count_bloat_types ,
       height = 3, width = 7,  units = c("in"), device = "pdf")

# List artefacts with more bloated declared and bloated undeclared dependencies
deps_count_types %>% filter(TypeUsage== "BloatedDeclared" & count_type_usage >100) %>% arrange(desc(count_type_usage))
deps_count_types %>% filter(TypeUsage== "BloatedUndeclared" & count_type_usage >100) %>% arrange(desc(count_type_usage))
  
```

#### Correlation between % of bloated deps and # transitive deps

```{r}
#DirectDeps/#TransitiveDeps vs #DeclaredBloated/#UndeclaredBloated
deps_count_types <- results %>%
  dplyr::filter(Scope == "compile" & Optional == F) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>%
  group_by(Artifact, TypeUsage) %>%
  summarise(n = n()) %>% 
   mutate(percentage = n / sum(n)*100) %>% 
   ungroup() %>% 
  tidyr::complete(Artifact, TypeUsage, fill = list(n = 0))

deps_count_types$percentage <- ifelse(is.na(deps_count_types$percentage), 0, deps_count_types$percentage)

NbBloatedDeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedDeclared") %>%
  dplyr::select(Artifact, n) %>% 
  dplyr::rename(NbBloatedDeclared = n)

NbBloatedUndeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedUndeclared") %>%
  dplyr::select(Artifact, n) %>% 
  dplyr::rename(NbBloatedUndeclared = n)

PercBloatedDeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedDeclared") %>%
  dplyr::select(Artifact, percentage) %>% 
  dplyr::rename(PercBloatedDeclared = percentage)

PercBloatedUnDeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedUndeclared") %>%
  dplyr::select(Artifact, percentage) %>% 
  dplyr::rename(PercBloatedUndeclared = percentage)

total_deps <- results %>% group_by(Artifact) %>% summarize(total_deps = n())

total_types <- results %>% group_by(Artifact, Type) %>% summarize(types = n())
NbDirectDeps <- total_types %>% 
  filter(Type == "direct") %>%
  dplyr::select(Artifact, n) %>% 
  dplyr::rename(NbBloatedDeclared = n)


matrix <- inner_join(inner_join(inner_join(inner_join(NbBloatedDeclared, total_deps, by = "Artifact"), NbBloatedUndeclared, by = "Artifact"), PercBloatedDeclared,  by = "Artifact"), PercBloatedUnDeclared, by = "Artifact") %>% select(-Artifact)

corr <- round(cor(matrix), 1)

library(ggcorrplot)

# Add correlation coefficients
# --------------------------------
# argument lab = TRUE
ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE)

```

#### Number of direct dependencies vs number of bloated dependencies

```{r}
results_types <- results_bloat %>% 
   group_by(Artifact, Type) %>% 
  summarise(count_types = n()) %>%
  filter(Type == "direct")

results_bloat_types <-  results %>% 
  dplyr::filter(Scope == "compile" & Optional == F) %>%
  mutate(TypeUsage = ifelse(UsedDeclared == T, "Used", ifelse(UsedUndeclared == T, "Used", ifelse(BloatedDeclared == T, "Bloated", "Bloated")))) %>% 
   group_by(Artifact, TypeUsage) %>% 
  summarise(count_type_usage = n()) 

join <- inner_join(results_types, results_bloat_types, by = "Artifact")

direct_bloated_point <- join %>% filter(TypeUsage == "Bloated") %>% 
  ggplot(aes(count_types, count_type_usage)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method='lm',formula=y~x) +
  labs(x = "#Direct dependencies", y = "#Bloated dependencies")
 
ggsave(filename = "Figures/direct_bloated_point.pdf", plot = direct_bloated_point ,
       height = 3, width = 7,  units = c("in"), device = "pdf")
```

Do the most popular libraries have more dependencies?

```{r}
data_usages <- data %>%
  dplyr::mutate(Artifact = paste(paste(GroupId, ArtifactId, sep = ":"), Version, sep = ":")) %>%
  dplyr::select(Artifact, DUsages)

results %>% dplyr::count(Artifact, Type)

bloat_ratio <- results_bloat %>% 
  dplyr::group_by(Artifact, TypeUsage) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(freq = n / sum(n)) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(Artifact) %>% 
  dplyr::mutate(TotalDeps= sum(n))

bloat_ratio <- bloat_ratio %>% filter(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared" )

bloat_popularity <- inner_join(bloat_ratio, data_usages)

library(corrplot)
M <- cor(bloat_popularity$freq, bloat_popularity$DUsages)


lm_eqn <- function(bloat_popularitTotalDeps){
    m <- lm(TotalDeps ~ freq, bloat_popularitTotalDeps);
    eq <- substitute(italic(TotalDeps) == a + b %.% italic(freq)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}

# ----------------------------------------------------------------------------
# correlation between bloated and number of dependencies
# ----------------------------------------------------------------------------
bloat_popularity %>% ggplot(aes(x = freq, y = TotalDeps)) +
  geom_smooth(method='lm', formula=y~x) +
  scale_fill_viridis(discrete = TRUE) +
  geom_point(alpha = 0.1) + 
  coord_cartesian( xlim = c(0, 1), ylim = c(0, 500)) +
  # add correlation labels
  # geom_text(x = 0.3, y = 300, label = lm_eqn(bloat_popularity), parse = TRUE) +
  labs(x="%Bloated dependencies", y="#Dependencies") +
  # scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "top", text = element_text(size = 16, family = "LM Roman 10"))

# geom_hex chart
bloat_deps_hex <-bloat_popularity %>% ggplot(aes(x = freq, y = TotalDeps)) +
  geom_hex() +
   scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  scale_fill_viridis() +
  labs(x="%Bloated dependencies", y="#Dependencies") +
  # scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "right", text = element_text(size = 16, family = "LM Roman 10"))

ggsave(filename = "Figures/bloat_deps_hex.pdf", plot = bloat_deps_hex,
       height = 3, width = 5,  units = c("in"), device = "pdf")

# ----------------------------------------------------------------------------
# correlation between bloated and direct usages
# ----------------------------------------------------------------------------
bloat_popularity %>% ggplot(aes(x = freq, y = DUsages)) +
  geom_smooth(method='lm', formula=y~x) +
  scale_fill_viridis(discrete = TRUE) +
  # add correlation labels
  # geom_text(x = 0.3, y = 300, label = lm_eqn(bloat_popularity), parse = TRUE) +
  labs(x="%Bloated dependencies", y="#Usages") +
  scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "top", text = element_text(size = 16, family = "LM Roman 10"))

# geom_hex chart
bloat_usages_hex <- bloat_popularity %>% ggplot(aes(x = freq, y = DUsages)) +
  geom_hex() +
   scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  scale_fill_viridis() +
  labs(x="%Bloated dependencies", y="#Usages") +
  scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "right", text = element_text(size = 16, family = "LM Roman 10"))

ggsave(filename = "Figures/bloat_usages_hex.pdf", plot = bloat_usages_hex,
       height = 3, width = 5,  units = c("in"), device = "pdf")

# ----------------------------------------------------------------------------
# ridgeline deep of the tree and percentage of bloated deps
# ----------------------------------------------------------------------------
bloat_treedeep<- inner_join(bloat_ratio, description, by = "Artifact") %>% dplyr::select(Artifact, TypeUsage, freq, HeightOriginalDT)
 
bloat_treedeep$HeightOriginalDT <- as.factor(bloat_treedeep$HeightOriginalDT)

# ridgeline with density plot
bloat_treedeep_ridges <- bloat_treedeep %>% 
  filter(is.na(HeightOriginalDT) == F) %>% 
  ggplot(aes(x = freq, y = HeightOriginalDT, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01,  stat="binline", bins=50) +
  scale_fill_viridis(name = "freq", option = "C") +
  # scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "right", text = element_text(size = 16, family = "LM Roman 10")) +
  labs(x=NULL, y="Height of the DT") +
  facet_wrap(~TypeUsage)

# ridgeline with histogram
bloat_treedeep_ridges <- bloat_treedeep %>% 
  filter(is.na(HeightOriginalDT) == F & TypeUsage == "BloatedUndeclared") %>%
  ggplot(aes(x = freq, y = HeightOriginalDT, fill = "#482677FF", color = "#482677FF")) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01,  stat="binline", bins=10) +
  # scale_fill_viridis(name = "freq", option = "C") +
  scale_x_continuous(range(0, 1)) +
  theme(legend.position = "none", text = element_text(size = 16, family = "LM Roman 10")) +
  labs(x=NULL, y="Depth of the dependency tree") 

ggsave(filename = "Figures/bloat_treedeep_ridges.pdf", plot = bloat_treedeep_ridges,
       height = 5, width = 4,  units = c("in"), device = "pdf") 

```



```{r}
results %>% filter()
colnames(results)
```


### Bloated dependencies in multi-module projects

```{r}
description_multimodule <- description %>% dplyr::select(Artifact, IsMultimodule)

# percentage of usage types for the compiled scope
filtered_results_bloat<- results_bloat %>% 
  filter(Scope %in% c("compile")) %>% 
  dplyr::group_by(Artifact, TypeUsage) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)) %>% 
  ungroup()

filtered_results_bloat$label <- round(filtered_results_bloat$percentage * 100, 1)

join_multimodule <- inner_join(filtered_results_bloat, description_multimodule, by = "Artifact") 

join_multimodule %>% 
  group_by(TypeUsage, IsMultimodule) %>% 
  summarise(sum(join_multimodule$n))

# count the number

join_multimodule_aggregated <- aggregate(join_multimodule$n, by=list(IsMultimodule=join_multimodule$IsMultimodule, TypeUsage=join_multimodule$TypeUsage), FUN=sum)

filtered_results_bloat_multimodule <- join_multimodule_aggregated %>% 
  ggplot(aes(x = IsMultimodule, y = x, fill = TypeUsage)) +
    geom_bar(stat="identity", position = "dodge") +
    # geom_text(size = 5, position = position_stack(vjust = 0.5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(x = "Multimodule", y = "#Dependencies") +
     scale_y_continuous(name="#Dependencies", labels = comma) +
    # scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip()

# count the percentage


join_multimodule_aggregated %>% group_by(IsMultimodule, TypeUsage) %>% summarise(x/sum(x))

join_multimodule_aggregated$IsMultimodule <- as.factor(join_multimodule_aggregated$IsMultimodule)
join_multimodule_aggregated$TypeUsage <- as.factor(join_multimodule_aggregated$TypeUsage)

join_multimodule_aggregated <- join_multimodule_aggregated %>% 
   dplyr::ungroup() %>% 
  group_by(IsMultimodule, TypeUsage) %>% 
  dplyr::mutate(freq = x / sum(x))
 

filtered_results_bloat_multimodule <- join_multimodule_aggregated %>% 
  ggplot(aes(x = IsMultimodule, y = x, fill = TypeUsage)) +
    geom_bar(stat="identity", position = "dodge") +
    # geom_text(size = 5, position = position_stack(vjust = 0.5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(x = "Multimodule", y = "#Dependencies") +
     scale_y_continuous(name="#Dependencies", labels = comma) +
    # scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip()

ggsave(filename = "Figures/filtered_results_bloat_multimodule.pdf", plot = filtered_results_bloat_multimodule,
       height = 3, width = 7,  units = c("in"), device = "pdf")
```


### Rotten bloated dependencies

```{r}
results %>% filter(AllDeps == "org.eclipse.core:filesystem:test")

View(results %>% group_by(Artifact) %>% summarise(n =n()) %>% arrange(desc(n)))

View(results %>% filter(Artifact == "org.jxls:jxls-poi:1.0.15"))

View(results %>% filter(BloatedDeclared == TRUE & Scope == "compile"))


results %>% filter(Artifact == "org.jxls:jxls-poi:1.0.15" & Scope == "compile") %>% filter(Used == F & Type == "transitive")
View(data %>% filter(GroupId == "org.apache.flink"))

results %>% group_by(Artifact) %>% summarise(n = n()) %>%  arrange(desc(n))

# org.apache.accumulo:accumulo-core:1.9.0
# org.apache.hive:hive-contrib:2.3.3
# org.apache.flink:flink-hbase_2.11:1.5.0
# org.apache.drill:drill-yarn:1.14.0
# com.netflix.dynomitemanager:dynomitemanager-common:2.0.13
# com.facebook.presto:presto-main:0.192
# ml.dmlc:xgboost4j-spark:0.72
# org.apache.hive.shims:hive-shims-0.23:2.3.3

filtered_results_bloat %>% filter(Artifact == "org.apache.accumulo:accumulo-core:1.9.0")

accumulo_versions_list <- data %>% 
  filter(GroupId == "org.apache.accumulo" & ArtifactId == "accumulo-core") %>% 
  arrange(desc(Version)) %>% 
  mutate(Artifact = paste(paste(GroupId, ArtifactId, sep = ":"), Version, sep = ":")) %>% 
  select(Artifact)

# write the list versions of the candidate library
write.csv(accumulo_versions_list, file = "/home/cesarsv/Documents/accumulo_versions_list.csv", col.names = FALSE, row.names = F, quote = F)

# read the dependency analysis results of all versions
accumulo_versions_list_results <- read_csv("Data/accumulo_versions_list_results.csv", col_names = T)

# reshape the dataset
accumulo_versions_list_results <- accumulo_versions_list_results %>%
  mutate(
    UsedDeclared =  ifelse(Used == T &
                             Declared == T, T, F),
    UsedUndeclared =  ifelse(Used == T &
                               Declared == F, T, F),
    BloatedDeclared =  ifelse(Used == F &
                                Declared == T, T, F),
    BloatedUndeclared =  ifelse(Used == F &
                                  Declared == F, T, F)
  ) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  group_by(Artifact, TypeUsage) %>% 
  summarise(n = n()) %>% 
  filter(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared")

# Extract release date for each version
data2 <- data %>% 
  filter(GroupId == "org.apache.accumulo" & ArtifactId == "accumulo-core") %>% 
  mutate(Artifact = paste(paste(GroupId, ArtifactId, sep = ":"), Version, sep = ":")) %>% 
  select(Artifact, Version, Release) 
  
accumulo_versions_list_results_plot <- inner_join(accumulo_versions_list_results, data2, by = "Artifact") %>% 
  ggplot(aes(x = Release, y = n, color = TypeUsage, shape = TypeUsage)) +
  geom_point(size = 3) +
  geom_line(aes(type = TypeUsage)) +
  geom_text(aes(label=Version), hjust=0, vjust=-0.5,show.legend = FALSE, size = 2) +
  xlab("Release Date") + 
  ylab("Count") + 
  # scale_color_viridis(discrete=TRUE) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",  limits = as.Date(c('2012-01-01','2019-01-01'))) +
  scale_y_continuous(limits=c(0, 100)) +
  scale_color_manual(values = c("#482173", "#2e6f8e")) +
  theme(legend.title = element_blank())

# viridis color palette: #482173, #2e6f8e, #29af7f, #bddf26
ggsave(filename = "Figures/accumulo_versions_list_results_plot.pdf", plot = accumulo_versions_list_results_plot,
       height = 3, width = 7,  units = c("in"), device = "pdf")
  
```




### How much dependency conflicts are due to bloated dependencies?

```{r}
dependency_status <- results %>% 
  dplyr::filter(Scope == "compile" & Optional == F) %>%
  mutate(TypeUsage = ifelse(UsedDeclared == T, "UsedDeclared", ifelse(UsedUndeclared == T, "UsedUndeclared", ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")))) %>%  
  dplyr::select(Artifact, InConflict, TypeUsage) %>% 
  group_by(Artifact, InConflict, TypeUsage) %>% 
  summarise(n = n()) 

dependency_status$InConflict <- ifelse(dependency_status$InConflict == "NO", "Other", ifelse(dependency_status$InConflict == "Ommited for duplicate", "Ommited for duplicate", "In conflict"))
  
dependency_status_barplot <- dependency_status %>% ggplot(aes(x = InConflict, y = n, fill=TypeUsage)) +
  geom_bar(stat="identity", position=position_dodge()) +
  xlab("Dependency Status") +
  ylab("#Dependencies") +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +

ggsave(filename = "Figures/dependency_status_barplot.pdf", plot = dependency_status_barplot,
       height = 3, width = 7,  units = c("in"), device = "pdf") 

```

### Comparison between the height of the original vs debloated dependency tree

```{r}
description$HeightOriginalDT <-  as.numeric(description$HeightOriginalDT)
description$HeightDebloatedDT <-  as.numeric(description$HeightDebloatedDT)

description %>% ggplot(aes(HeightOriginalDT)) +
  geom_histogram()

description %>% ggplot(aes(HeightDebloatedDT)) +
  geom_histogram()

metrics %>%
  filter(is.na(isPassive) == F) %>%
  mutate(Type = ifelse(isPassive == "true", "passive", "active")) %>% 
  # filter(Type == "passive") %>% 
  # filter(validityPeriod > 0) %>% 
  ggplot(aes(Type, lifespan)) +
  geom_violin(trim = FALSE, fill = "#CCCCCC") +
  geom_boxplot(width = 0.1) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x)
      10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  coord_flip() +
  theme(legend.position = "none") +
  xlab(NULL) +
  ylab(TeX("lifespan (log_{10} scale)"))
  
```


```{r}
# library
library(ggridges)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/zonination/perceptions/master/probly.csv", header=TRUE, sep=",")
data <- data %>% 
  gather(key="text", value="value") %>%
  mutate(text = gsub("\\.", " ",text)) %>%
  mutate(value = round(as.numeric(value),0)) %>%
  filter(text %in% c("Almost Certainly","Very Good Chance","We Believe","Likely","About Even", "Little Chance", "Chances Are Slight", "Almost No Chance"))

# Plot
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(y=text, x=value,  fill=text)) +
    geom_density_ridges(alpha=0.6, stat="binline", bins=20) +
    theme_ridges() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```




```{r}
# data preparation library(ggplot2)
col <- c("red", "pink", "orange", "brown")
freq <- c(101, 68, 34, 18)

tmp %>% group_by(Artifact,Type) %>% 
  summarise(count=n())
## create data frame
colour.df <- data.frame(col, freq)
colour.df

## calculate percentage 
colour.df$percentage = colour.df$freq / sum(colour.df$freq)* 100
colour.df = colour.df[rev(order(colour.df$percentage)), ]
colour.df$ymax = cumsum(colour.df$percentage)
colour.df$ymin = c(0, head(colour.df$ymax, n = -1))
colour.df

library(ggrepel)
colour.df %>% ggplot(aes(fill = col, ymax = ymax, ymin = ymin, xmax = 100, xmin = 80)) +
    geom_rect(colour = "black") +
    coord_polar(theta = "y") + 
    xlim(c(0, 100)) +
    geom_label_repel(aes(label = paste(round(percentage,2),"%"), x = 100, y = (ymin + ymax)/2),inherit.aes = F, show.legend = F, size = 5)+
    theme(legend.title = element_text(colour = "black", size = 16, face = "bold"), 
        legend.text = element_text(colour = "black", size = 15), 
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()) +
    annotate("text", x = 0, y = 0, size = 8, label = "Dependencies") +
  theme(legend.position = "top",
        text = element_text(size = 16, family = "LM Roman 10"))
    
```

### Plot conflicts

```{r}
results <-
  results %>% 
  filter(Type == "transitive") %>% 
  mutate(UsageType = ifelse(
    UsedDeclared == "true",
    "UsedDeclared",
    ifelse(
      UsedUndeclared == "true",
      "UsedUndeclared",
      ifelse(BloatedDeclared == "true", "BloatedDeclared", "BloatedUndeclared")
    )
  )) 

results %>% filter(UsageType == "UsedUndeclared" | UsageType == "BloatedUndeclared") %>% 
  ggplot(aes(UsageType, ..count..)) + geom_bar(aes(fill = InConflict), position = "dodge") +
   theme(legend.position = "top", text=element_text(size=14, family="LM Roman 10"))

ggsave(filename = "Figures/usage-conflicts.pdf", height = 4, width = 6,  units = c("in"), device = cairo_pdf)

```