---
title: "Data analysis to answer the research questions"
output: html_notebook
---

Import required data and libraries

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# clear the R environment
rm(list=ls())

library(tidyverse)
library(latex2exp)
library(scales)
library(extrafont)
library(viridis)
library(ggridges)
library(forcats)
library(ggExtra)
library(grid)
library(gtable)
# set fonts
loadfonts(device = "pdf")
# font_import()
# link www.fontsquirrel.com/fonts/latin-modern-roman  

# execute once to add fonts:
# font_import(pattern = "lmroman*")
# theme(legend.position = "top", text=element_text(size=14, family="LM Roman 10"))

# set ggplot global theme
theme_set(theme_bw() +
            theme(legend.position = "top") +
            theme(text = element_text(size = 16, family = "LM Roman 10")))

# multiple figures together
if(!require(devtools)) install.packages("devtools")
devtools::install_github("thomasp85/patchwork")
library(patchwork)

# load pre-processed data
load(file="Data/description.RData")
load(file="Data/results.RData")
data <- read_csv("Data/data.csv", col_names = T)

```

# Transitive dependency explosion

## Descriptive statistics of direct and transitive dependencies

```{r}
nb_direct_and_transitive_deps <- 
  results %>%
  dplyr::group_by(Artifact, Type) %>%
  dplyr::summarize(n = n()) %>%
   mutate(percentage = n / sum(n)*100) %>% 
  ungroup() %>% 
  tidyr::complete(Artifact, Type, fill = list(n = 0))

# summary statistics of direct and transitive dependencies
direct_deps <- nb_direct_and_transitive_deps %>% filter(Type == "direct")
transitive_deps <- nb_direct_and_transitive_deps  %>% filter(Type == "transitive")
summary(direct_deps$n)
summary(transitive_deps$n)

print("percentage of direct deps:")
(sum(direct_deps$n) * 100) / (sum(direct_deps$n) + sum(transitive_deps$n))
print("percentage of transitive deps:")
(sum(transitive_deps$n) * 100) / (sum(direct_deps$n) + sum(transitive_deps$n))

nb_direct_and_transitive_deps$Type <- ifelse(nb_direct_and_transitive_deps$Type == "direct", "Direct", "Transitive")

# save the data objects
save(nb_direct_and_transitive_deps, file = "Data/nb_direct_and_transitive_deps.RData")
load(file = "Data/nb_direct_and_transitive_deps.RData")
```

## Ratio of bloat vs ratio of direct/transitive

```{r}

```

## Correlation between direct and transitive dependencies

```{r}
direct_n <- nb_direct_and_transitive_deps %>% filter(Type == "Direct") %>% dplyr::rename(Direct = n)
transitive_n <- nb_direct_and_transitive_deps %>% filter(Type == "Transitive") %>% dplyr::rename(Transitive = n)

p <- inner_join(direct_n, transitive_n, by = "Artifact") %>% 
  ggplot(aes(x = Direct, y = Transitive)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method='lm',formula=y~x) +
   scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  labs(x = "#Direct dependencies", y = "#Transitive dependencies")

p <- ggExtra::ggMarginal(p, type = "boxplot") 

ggsave(filename = "Figures/nb_direct_and_transitive_deps_boxplot.pdf", plot = p,
       height = 3, width = 7,  units = c("in"), device = "pdf")
```

## Violinplot with the distribution of direct and transitive dependencies

```{r}
# plot boxplot the distribution of direct and transitive dependencies
means <- aggregate(n ~  Type, nb_direct_and_transitive_deps, mean)
means$n <- round(means$n, 2)
nb_direct_and_transitive_deps_boxplot <- 
  nb_direct_and_transitive_deps %>% 
  ggplot(aes(Type, n)) +
  geom_violin(trim = FALSE, aes(fill = Type)) +
  geom_boxplot(width = 0.1) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  geom_point(data = means,
             aes(x=Type, y=n),
             colour="black",
                shape=18, size=6, show.legend = FALSE)+
  geom_point(data = means,
             aes(x=Type, y=n),
             colour="lightblue", 
                shape=18, size=4, show.legend = FALSE)+
  geom_text(data = means, aes(label = n, y = n)) +
  scale_fill_viridis(discrete = TRUE, option = "D", alpha = 0.75) +
  ylab(TeX("#Dependencies")) +
  xlab("Type") +
  coord_flip() +
  theme(legend.position = "none")

ggsave(filename = "Figures/nb_direct_and_transitive_deps_boxplot.pdf", plot = nb_direct_and_transitive_deps_boxplot,
       height = 3, width = 7,  units = c("in"), device = "pdf")

```

## Geom_area with the distribution of direct and transitive dependencies

```{r}

View(nb_direct_and_transitive_deps)

# case transitive == zero
nb_direct_and_transitive_deps$percentage <- ifelse(is.na(nb_direct_and_transitive_deps$percentage), 0, nb_direct_and_transitive_deps$percentage)

nb_direct_and_transitive_deps$Artifact <- as.numeric(as.factor(nb_direct_and_transitive_deps$Artifact))
nb_direct_and_transitive_deps$Artifact <- as.factor(nb_direct_and_transitive_deps$Artifact)

nb_direct_and_transitive_deps_arranged <- nb_direct_and_transitive_deps %>% 
  group_by(Artifact, Type) %>% 
  arrange(percentage)

nb_direct_and_transitive_deps_arranged$Artifact <- as.numeric(nb_direct_and_transitive_deps_arranged$Artifact)

nb_direct_and_transitive_deps_arranged_direct <- nb_direct_and_transitive_deps_arranged %>% filter(Type == "Direct")
nb_direct_and_transitive_deps_arranged_transitive <- nb_direct_and_transitive_deps_arranged %>% filter(Type == "Transitive")

nb_direct_and_transitive_deps_arranged_direct$Artifact <- c(1:10000)
nb_direct_and_transitive_deps_arranged_transitive$Artifact <- c(10000:1)

nb_direct_and_transitive_deps_arranged <- bind_rows(nb_direct_and_transitive_deps_arranged_direct, nb_direct_and_transitive_deps_arranged_transitive)

my3cols <- c("#E7B800", "#2E9FDF")
nb_direct_and_transitive_deps_area <-
  nb_direct_and_transitive_deps_arranged %>%
  ggplot(aes(
    x = Artifact, 
    y = percentage,
    fill = Type)) +
  geom_area(alpha = 0.9, position = "fill") +
  ylab("%Dependencies") +
  scale_x_discrete() +
  xlab("Projects") +
  xlim(0, 10000) +
  scale_y_continuous(labels = scales::percent) +
  scale_fill_manual(values = my3cols) +
  theme( legend.title = element_blank())

# deps_area_deps_boxplot <- nb_direct_and_transitive_deps_area + nb_direct_and_transitive_deps_boxplot +plot_layout(ncol = 1)

ggsave(filename = "Figures/nb_direct_and_transitive_deps_area.pdf", plot = nb_direct_and_transitive_deps_area,
       height = 2.5, width = 7,  units = c("in"), device = "pdf")

```

# How frequent is the occurrence of bloated dependencies in the Maven ecosystem?

## Boxplot with the Types of Usages

```{r}
results_bloat <- results %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  mutate(TypeUsage = ifelse(UsedDeclared == T, "UsedDeclared", ifelse(UsedUndeclared == T, "UsedUndeclared", ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")))) %>% 
  dplyr::select(Artifact, TypeUsage) %>% 
   dplyr::group_by(Artifact, TypeUsage) %>%
  dplyr::summarize(n = n()) %>%
  ungroup() %>% 
  tidyr::complete(Artifact, TypeUsage, fill = list(n = 0))

# calculate the means
means <- aggregate(n ~  TypeUsage, results_bloat, mean)
means$n <- round(means$n, 2)

filtered_type_bloat_boxplot <- results_bloat %>% 
  ggplot(aes(TypeUsage, n, fill = TypeUsage)) +
  geom_boxplot() +
  geom_point(data = means,
             aes(x=TypeUsage, y=n),
             colour="black",
                shape=18, size=6, show.legend = FALSE)+
  geom_point(data = means,
             aes(x=TypeUsage, y=n),
             colour="lightblue", 
                shape=18, size=4, show.legend = FALSE)+
  geom_text(data = means, aes(label = n, y = n + 2)) +
  scale_y_continuous(breaks=c(0, 1,10,100,1000), trans="log1p") +
   # scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
   #            labels = trans_format("log10", math_format(10^.x))) +
  xlab (NULL) +
  ylab ("#Dependencies") +
  # annotation_logticks(sides="l") +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10"))  +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

# show the summary
BD <- results_bloat %>% filter(TypeUsage == "BloatedDeclared")
BU <- results_bloat %>% filter(TypeUsage == "BloatedUndeclared")
UU <- results_bloat %>% filter(TypeUsage == "UsedUndeclared")
UD <- results_bloat %>% filter(TypeUsage == "UsedDeclared")
print("BloatedDeclared")
summary(BD$n)
print("BloatedUndeclared")
summary(BU$n)
print("UsedUndeclared")
summary(UU$n)
print("UsedDeclared")
summary(UD$n)

ggsave(filename = "Figures/filtered_type_bloat_boxplot.pdf", plot = filtered_type_bloat_boxplot ,
       height = 4, width = 7,  units = c("in"), device = "pdf")

```

## Count of projects with all their dependencies declared

```{r}
library(stringr)

a <- results %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  select(Artifact, AllDeps, Declared, Used) %>%
  group_by(Artifact, Declared, Used) %>% 
  summarise(count = n()) %>% 
  ungroup() %>% 
  tidyr::complete(Artifact, Declared, Used, fill = list(count = 0))

b <- results %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  select(Artifact, AllDeps) %>% 
  group_by(Artifact) %>% 
  summarise(total = n()) 

inner_join(a, b, by= "Artifact") %>% 
  filter(Declared == TRUE & Used == TRUE & count == total)
  
# dependencies that are always used and declared
a <- results %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  select(Artifact, AllDeps, Declared, Used) %>% 
  filter(Declared == TRUE & Used == TRUE) %>% 
  select(AllDeps) %>% 
  distinct()


a <- a %>% 
  separate(AllDeps, c("GroupID", "ArtifactId", "Version"), ":") %>% 
  mutate(Lib = paste(GroupID, ArtifactId, sep = ":")) %>% 
  select(Lib) %>% 
  distinct()

b <- results %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  select(Artifact, AllDeps, Declared, Used) %>% 
  filter(Declared != TRUE | Used != TRUE) %>% 
  select(AllDeps) %>% 
  distinct()

b <- b %>% 
  separate(AllDeps, c("GroupID", "ArtifactId", "Version"), ":") %>% 
  mutate(Lib = paste(GroupID, ArtifactId, sep = ":")) %>% 
  select(Lib) %>% 
  distinct()

c <- setdiff(a, b)

x <- results %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  select(AllDeps) %>% 
  separate(AllDeps, c("GroupID", "ArtifactId", "Version"), ":") %>% 
  mutate(Lib = paste(GroupID, ArtifactId, sep = ":")) %>% 
  group_by(Lib) %>% 
  summarise(n = n())

View(inner_join(x,c, by= "Lib") %>% 
  arrange(desc(n)))




results %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  select(AllDeps, Declared, Used) %>% 
  group_by(AllDeps, Declared, Used) %>% 
  summarise(count=n()) %>% 
  ungroup() %>% 
  tidyr::complete(AllDeps, Declared, Used, fill = list(count = 0))
  
  
```

## Barplot with the Types of Usages

```{r}
results_bloat <- results %>% 
  # filter(Scope == "compile", Optional == "false", Type == "direct") %>% 
  mutate(TypeUsage = ifelse(UsedDeclared == T, "UsedDeclared", ifelse(UsedUndeclared == T, "UsedUndeclared", ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared"))))

# percentage of each scope
percentage_scopes <- results_bloat %>% 
  filter(Scope %in% c("compile", "test", "runtime", "provided")) %>% 
  dplyr::group_by(Scope) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)*100) %>% 
  ungroup()

# percentage of usage types for the compiled scope
filtered_results_bloat<- results_bloat %>% 
  filter(Scope %in% c("compile")) %>% 
  dplyr::group_by(TypeUsage) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)) %>% 
  ungroup() 

filtered_results_bloat$label <- round(filtered_results_bloat$percentage * 100, 1)
 
filtered_type_bloat_barplot <- filtered_results_bloat %>% 
  ggplot(aes(x = "Compile", y = percentage, fill = TypeUsage, label = paste0(label,"%"))) +
    geom_bar(stat="identity", position = position_fill(reverse = TRUE)) +
      geom_text(aes(y = percentage, label = paste0(label,"%")), position = position_stack(reverse = TRUE, vjust = .5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(y = "%Dependencies", x = "Scope") +
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip()

ggsave(filename = "Figures/filtered_type_bloat_barplot.pdf", plot = filtered_type_bloat_barplot ,
       height = 2.5, width = 7,  units = c("in"), device = "pdf")

# Examples of popular libraries with many bloated dependencies
results_bloat %>% group_by(Artifact, TypeUsage) %>% summarise(n = n()) %>% arrange(desc(n)) %>% 
  filter(Artifact == "org.wso2.carbon:org.wso2.carbon.core:4.4.27")

results_bloat %>% group_by(Artifact, TypeUsage) %>% summarise(n = n()) %>% arrange(desc(n)) %>% 
  filter(Artifact == "org.eclipse.che.core:che-core-api-core:6.10.0")
```

## Set of unique dependencies

```{r}

tmp <- results_bloat %>% distinct(AllDeps, .keep_all= TRUE) %>% filter(Type == "direct") 

table(tmp$TypeUsage)

```


## Impact of bloat per community

```{r}
description %>% count(Organization) %>% arrange(desc(n))

# filter Apache Software Foundation projects
description %>% filter(str_detect(Artifact, "org.apache"))

# filter Eclipse Foundation artifacts
description %>% filter(str_detect(Artifact, "org.eclipse"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "org.wso2"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.facebook"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.amazon"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.netflix"))

# filter Twitter Inc. projects
description %>% filter(str_detect(Artifact, "com.google"))

communities <- results_bloat %>% select(Artifact, AllDeps, TypeUsage) %>% mutate(Community =
                                                           ifelse(
                                                             startsWith(Artifact, "org.apache"),
                                                             "Apache (1103)",
                                                              ifelse(
                                                               startsWith(Artifact, "org.eclipse"),
                                                               "Eclipse (406)", 
                                                               ifelse(startsWith(Artifact, "org.wso2"),
                                                                      "OW2 (536)", 
                                                                      ifelse(
                                                                        startsWith(Artifact, "com.facebook"),
                                                                                   "Facebook (17)", 
                                                                                   ifelse(
                                                                                     startsWith(Artifact, "com.amazon"),
                                                                                                "Amazon (91)",
                                                                                                ifelse(
                                                                                                  startsWith(Artifact, "com.netflix"),
                                                                                                             "Netflix (51)", 
                                                                                                             ifelse(
                                                                                                               startsWith(Artifact, "com.google"), "Google (62)", "Other")
                                                                                                             ))
                                                                                                ))
                                                                                   ))
                                                                      )
# percentage of usage types for per community
communities <- communities %>% 
   filter(Community != "OW2 (536)" & Community != "Other") %>% 
  dplyr::group_by(Community, TypeUsage) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)) %>% 
  ungroup() 

communities$label <- round(communities$percentage * 100, 1)

# communities$TypeUsage <- factor(communities$TypeUsage, levels = rev(levels(communities$TypeUsage)))

communities$Community <- factor(communities$Community,levels = c("Google (62)", "Netflix (51)", "Amazon (91)", "Facebook (17)", "Eclipse (406)", "Apache (1103)" ))

communities_bloat <- communities %>% 
  ggplot(aes(x = Community, y = percentage, fill = TypeUsage, label = paste0(label,"%"))) +
    geom_bar(stat="identity", position = position_fill(reverse = TRUE)) +
    geom_text(aes(y = percentage, label = paste0(label,"%")), position = position_stack(reverse = TRUE, vjust = .5), family = "LM Roman 10") +
    # geom_text(size = 4, position = position_stack(vjust = 0.5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(y = "%Dependencies", x = "Communities") +
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip() 

ggsave(filename = "Figures/communities_bloat.pdf", plot = communities_bloat ,
       height = 3, width = 8,  units = c("in"), device = "pdf")
                                                          
```

## Distribution of bloated dependencies across dependencies

```{r}
deps_count_types <- results %>%
  dplyr::filter(Scope == "compile" & Optional == F) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>%
  group_by(Artifact, TypeUsage) %>%
  summarise(count_type_usage = n())

total_deps <- results %>% group_by(Artifact) %>% summarize(total_deps = n())

dist_deps_count_bloat_types <- inner_join(deps_count_types, total_deps, by = "Artifact") %>% 
  filter(TypeUsage== "BloatedDeclared" | TypeUsage== "BloatedUndeclared") %>%
  ggplot(aes(x = count_type_usage, y=total_deps, color = TypeUsage, shape = TypeUsage)) +
  geom_point(alpha = 0.5, size = 3) +
  geom_smooth(se=F) +
  # scale_y_log10(
  #   breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
  #   labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  # ) +
  scale_y_continuous(limits=c(0, 5000)) +
  xlab("#Bloated dependencies") +
  ylab("#Dependencies") +
  # scale_color_viridis(discrete = TRUE) +
  scale_color_manual(values = c("#482173", "#2e6f8e")) +
  theme(legend.title = element_blank())

ggsave(filename = "Figures/dist_deps_count_bloat_types.pdf", plot = dist_deps_count_bloat_types ,
       height = 3, width = 7,  units = c("in"), device = "pdf")

# List artefacts with more bloated declared and bloated undeclared dependencies
deps_count_types %>% filter(TypeUsage== "BloatedDeclared" & count_type_usage >100) %>% arrange(desc(count_type_usage))
deps_count_types %>% filter(TypeUsage== "BloatedUndeclared" & count_type_usage >100) %>% arrange(desc(count_type_usage))
  
```

## Correlation between % of bloated deps and # transitive deps

```{r}
library(ggcorrplot)

#DirectDeps/#TransitiveDeps vs #DeclaredBloated/#UndeclaredBloated
deps_count_types <- results %>%
  dplyr::filter(Scope == "compile" & Optional == F) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>%
  group_by(Artifact, TypeUsage) %>%
  summarise(n = n()) %>% 
   mutate(percentage = n / sum(n)*100) %>% 
   ungroup() %>% 
  tidyr::complete(Artifact, TypeUsage, fill = list(n = 0))

deps_count_types$percentage <- ifelse(is.na(deps_count_types$percentage), 0, deps_count_types$percentage)

NbBloatedDeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedDeclared") %>%
  dplyr::select(Artifact, n) %>% 
  dplyr::rename(NbBloatedDeclared = n)

NbBloatedUndeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedUndeclared") %>%
  dplyr::select(Artifact, n) %>% 
  dplyr::rename(NbBloatedUndeclared = n)

PercBloatedDeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedDeclared") %>%
  dplyr::select(Artifact, percentage) %>% 
  dplyr::rename(PercBloatedDeclared = percentage)

PercBloatedUnDeclared <- deps_count_types %>% 
  filter(TypeUsage == "BloatedUndeclared") %>%
  dplyr::select(Artifact, percentage) %>% 
  dplyr::rename(PercBloatedUndeclared = percentage)

total_deps <- results %>% group_by(Artifact) %>% summarize(total_deps = n())

total_types <- results %>% group_by(Artifact, Type) %>% summarize(types = n())
NbDirectDeps <- total_types %>% 
  filter(Type == "direct") %>%
  dplyr::select(Artifact, types) %>% 
  dplyr::rename(NbDirectDeps = types)

NbTransitiveDeps <- total_types %>% 
  filter(Type == "transitive") %>%
  dplyr::select(Artifact, types) %>% 
  dplyr::rename(NbTransitiveDeps = types)

matrix <- inner_join(inner_join(inner_join(inner_join(inner_join(inner_join(NbBloatedDeclared, total_deps, by = "Artifact"), NbBloatedUndeclared, by = "Artifact"), PercBloatedDeclared,  by = "Artifact"), PercBloatedUnDeclared, by = "Artifact"), NbDirectDeps, by = "Artifact"), NbTransitiveDeps, by = "Artifact") %>% select(-c(Artifact, total_deps))

corr <- round(cor(matrix), 4)

# matrix of p values
p.mat <- cor_pmat(matrix)
View(as.data.frame(p.mat))
ggcorrplot(p.mat, hc.order = TRUE,
    type = "lower", p.mat = p.mat, sig.level = 0.001)

# Add correlation coefficients
# --------------------------------
# argument lab = TRUE

corrplot <- ggcorrplot(corr, hc.order = TRUE, type = "lower", legend.title = "Correlation",
   lab = TRUE,  p.mat = p.mat) +
  xlab(NULL) +
   ylab(NULL) +
  theme_set(theme_bw() +
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
             theme(text = element_text(size = 16, family = "LM Roman 10")))

ggsave(filename = "Figures/corrplot.pdf", plot = corrplot ,
       height = 6, width = 6,  units = c("in"), device = "pdf")

```

## Number of direct dependencies vs number of bloated dependencies

```{r}
results_types <- results_bloat %>% 
   group_by(Artifact, Type) %>% 
  summarise(count_types = n()) %>%
  filter(Type == "direct")

results_bloat_types <-  results %>% 
  dplyr::filter(Scope == "compile" & Optional == F) %>%
  mutate(TypeUsage = ifelse(UsedDeclared == T, "Used", ifelse(UsedUndeclared == T, "Used", ifelse(BloatedDeclared == T, "Bloated", "Bloated")))) %>% 
   group_by(Artifact, TypeUsage) %>% 
  summarise(count_type_usage = n()) 

join <- inner_join(results_types, results_bloat_types, by = "Artifact")

direct_bloated_point <- join %>% filter(TypeUsage == "Bloated") %>% 
  ggplot(aes(count_types, count_type_usage)) +
  geom_point(alpha = 0.2) +
  geom_smooth(method='lm',formula=y~x) +
  labs(x = "#Direct dependencies", y = "#Bloated dependencies")
 
ggsave(filename = "Figures/direct_bloated_point.pdf", plot = direct_bloated_point ,
       height = 3, width = 7,  units = c("in"), device = "pdf")
```

## Correlation between library popularity and bloated dependencies

```{r}
data_usages <- data %>%
  dplyr::mutate(Artifact = paste(paste(GroupId, ArtifactId, sep = ":"), Version, sep = ":")) %>%
  dplyr::select(Artifact, DUsages)

results_bloat <- results %>%
  filter(Scope == "compile", Optional == FALSE) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == TRUE,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == TRUE,
      "UsedUndeclared",
      ifelse(
        BloatedDeclared == TRUE,
        "BloatedDeclared",
        "BloatedUndeclared"
      )
    )
  )) %>%
  dplyr::count(Artifact, TypeUsage) %>% 
  ungroup() %>% 
  tidyr::complete(Artifact, TypeUsage, fill = list(n = 0))

join <- inner_join(data_usages, results_bloat, by = "Artifact")

join %>%
  ggplot(aes(x = DUsages, y = n)) +
  # geom_point() +
  stat_density_2d(aes(fill = stat(nlevel)), geom = "polygon") +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x)
      10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  scale_x_log10(
    breaks = scales::trans_breaks("log10", function(x)
      10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  facet_grid(TypeUsage ~ .) +
  scale_fill_viridis() +
  theme(legend.position = "right") 

#  --------------------- frequency
results_bloat <- results %>%
  filter(Scope == "compile", Optional == FALSE) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == TRUE,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == TRUE,
      "UsedUndeclared",
      ifelse(
        BloatedDeclared == TRUE,
        "BloatedDeclared",
        "BloatedUndeclared"
      )
    )
  )) %>%
  dplyr::count(Artifact, TypeUsage) %>% 
  ungroup() %>% 
  tidyr::complete(Artifact, TypeUsage, fill = list(n = 0)) %>% 
  dplyr::group_by(Artifact) %>% 
  dplyr::mutate(TotalDeps= sum(n)) %>% 
   dplyr::mutate(freq = n / TotalDeps)


join <- inner_join(data_usages, results_bloat , by = "Artifact")
join %>%
  filter(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared") %>% 
  ggplot(aes(x = freq, y = DUsages)) +
  #  scale_y_log10(
  #   breaks = scales::trans_breaks("log10", function(x)
  #     10 ^ x),
  #   labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  # ) +
  # scale_x_log10(
  #   breaks = scales::trans_breaks("log10", function(x)
  #     10 ^ x),
  #   labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  # ) +
  xlab("#Dependencies") +
  ylab("#Usages") +
  geom_point(alpha = 0.1) +
  facet_grid(. ~ TypeUsage)




results_bloat <- results %>%
  filter(Scope == "compile", Optional == FALSE) %>%
  mutate(IsBloated = ifelse(
    UsedDeclared == TRUE,
    FALSE,
    ifelse(
      UsedUndeclared == TRUE,
      FALSE,
      ifelse(
        BloatedDeclared == TRUE,
        TRUE,
        TRUE
      )
    )
  )) %>%
  dplyr::count(Artifact, IsBloated) %>% 
  ungroup() %>% 
  tidyr::complete(Artifact, IsBloated, fill = list(n = 0)) %>% 
  dplyr::group_by(Artifact) %>% 
  dplyr::mutate(TotalDeps= sum(n)) %>% 
   dplyr::mutate(PercBloat = n / TotalDeps)

perc_bloat <- results_bloat %>% 
  filter(IsBloated == TRUE) %>% 
  select(Artifact, PercBloat)

nb_direct_and_transitive_deps <- 
  results %>%
  filter(Scope == "compile", Optional == FALSE) %>%
  dplyr::group_by(Artifact, Type) %>%
  dplyr::summarize(n = n()) %>%
   mutate(PercType = n / sum(n)*100) %>% 
  ungroup() %>% 
  tidyr::complete(Artifact, Type, fill = list(n = 0)) %>% 
  select(Artifact, Type, PercType)


direct_bloated <- inner_join(perc_bloat, nb_direct_and_transitive_deps, by = "Artifact") %>% 
   mutate(PercType = PercType/100) %>% 
  filter(Type == "direct") %>% 
  ggplot(aes(x = PercType, y=PercBloat)) +
  geom_point(alpha = 0.1, color = "#E7B800") +
  geom_density_2d(color = "black") +
  xlab("%Direct Dependencies") +
  ylab("%Bloated Dependencies") +
  scale_x_continuous(labels = scales::percent) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "none")

transitive_bloated <- inner_join(perc_bloat, nb_direct_and_transitive_deps, by = "Artifact") %>% 
   mutate(PercType = PercType/100) %>% 
  filter(Type == "transitive") %>% 
  ggplot(aes(x = PercType, y=PercBloat)) +
  geom_point(alpha = 0.1, color = "#2E9FDF") +
  geom_density_2d(color = "black") +
  xlab("%Transitive dependencies") +
  ylab("%Bloated dependencies") +
  scale_x_continuous(labels = scales::percent, limits=c(0, 1)) +
  scale_y_continuous(labels = scales::percent) +
  theme(legend.position = "none")

dt_bloated_percentage <- direct_bloated + transitive_bloated

ggsave(filename = "Figures/transitive_bloated.pdf", plot = transitive_bloated,
       height = 4, width = 7,  units = c("in"), device = "pdf")

```

## Do the most popular libraries have more dependencies?

```{r}
data_usages <- data %>%
  dplyr::mutate(Artifact = paste(paste(GroupId, ArtifactId, sep = ":"), Version, sep = ":")) %>%
  dplyr::select(Artifact, DUsages)

results %>% dplyr::count(Artifact, Type)

bloat_ratio <- results_bloat %>% 
  dplyr::group_by(Artifact, TypeUsage) %>%
  dplyr::summarise (n = n()) %>%
  dplyr::mutate(freq = n / sum(n)) %>% 
  dplyr::ungroup() %>% 
  dplyr::group_by(Artifact) %>% 
  dplyr::mutate(TotalDeps= sum(n))

bloat_ratio <- bloat_ratio %>% filter(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared" )

bloat_popularity <- inner_join(bloat_ratio, data_usages)

library(corrplot)
M <- cor(bloat_popularity$freq, bloat_popularity$DUsages)

lm_eqn <- function(bloat_popularitTotalDeps){
    m <- lm(TotalDeps ~ freq, bloat_popularitTotalDeps);
    eq <- substitute(italic(TotalDeps) == a + b %.% italic(freq)*","~~italic(r)^2~"="~r2, 
         list(a = format(unname(coef(m)[1]), digits = 2),
              b = format(unname(coef(m)[2]), digits = 2),
             r2 = format(summary(m)$r.squared, digits = 3)))
    as.character(as.expression(eq));
}

# ----------------------------------------------------------------------------
# correlation between bloated and number of dependencies
# ----------------------------------------------------------------------------
bloat_popularity %>% ggplot(aes(x = freq, y = TotalDeps)) +
  geom_smooth(method='lm', formula=y~x) +
  scale_fill_viridis(discrete = TRUE) +
  geom_point(alpha = 0.1) + 
  coord_cartesian( xlim = c(0, 1), ylim = c(0, 500)) +
  # add correlation labels
  # geom_text(x = 0.3, y = 300, label = lm_eqn(bloat_popularity), parse = TRUE) +
  labs(x="%Bloated dependencies", y="#Dependencies") +
  # scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "top", text = element_text(size = 16, family = "LM Roman 10"))

# geom_hex chart
bloat_deps_hex <-bloat_popularity %>% ggplot(aes(x = freq, y = TotalDeps)) +
  geom_hex() +
   scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  scale_fill_viridis() +
  labs(x="%Bloated dependencies", y="#Dependencies") +
  # scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "right", text = element_text(size = 16, family = "LM Roman 10"))

ggsave(filename = "Figures/bloat_deps_hex.pdf", plot = bloat_deps_hex,
       height = 3, width = 5,  units = c("in"), device = "pdf")

# ----------------------------------------------------------------------------
# correlation between bloated and direct usages
# ----------------------------------------------------------------------------
bloat_popularity %>% ggplot(aes(x = freq, y = DUsages)) +
  geom_smooth(method='lm', formula=y~x) +
  scale_fill_viridis(discrete = TRUE) +
  geom_point() +
  # add correlation labels
  # geom_text(x = 0.3, y = 300, label = lm_eqn(bloat_popularity), parse = TRUE) +
  labs(x="%Bloated dependencies", y="#Usages") +
  # scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "top", text = element_text(size = 16, family = "LM Roman 10"))

# geom_hex chart
bloat_usages_hex <- bloat_popularity %>% ggplot(aes(x = freq, y = DUsages)) +
  geom_hex() +
   scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  scale_fill_viridis() +
  labs(x="%Bloated dependencies", y="#Usages") +
  scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "right", text = element_text(size = 16, family = "LM Roman 10"))

ggsave(filename = "Figures/bloat_usages_hex.pdf", plot = bloat_usages_hex,
       height = 3, width = 5,  units = c("in"), device = "pdf")

# ----------------------------------------------------------------------------
# ridgeline deep of the tree and percentage of bloated deps
# ----------------------------------------------------------------------------
bloat_treedeep<- inner_join(bloat_ratio, description, by = "Artifact") %>% dplyr::select(Artifact, TypeUsage, freq, HeightOriginalDT)
 
bloat_treedeep$HeightOriginalDT <- as.factor(bloat_treedeep$HeightOriginalDT)

# ridgeline with density plot
bloat_treedeep_ridges <- bloat_treedeep %>% 
  filter(is.na(HeightOriginalDT) == F) %>% 
  ggplot(aes(x = freq, y = HeightOriginalDT, fill = ..x..)) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01,  stat="binline", bins=50) +
  scale_fill_viridis(name = "freq", option = "C") +
  # scale_x_continuous(labels = scales::percentage) +
  theme(legend.position = "right", text = element_text(size = 16, family = "LM Roman 10")) +
  labs(x=NULL, y="Height of the DT") +
  facet_wrap(~TypeUsage)

# ridgeline with histogram
bloat_treedeep_ridges <- bloat_treedeep %>% 
  filter(is.na(HeightOriginalDT) == F & TypeUsage == "BloatedUndeclared") %>%
  ggplot(aes(x = freq, y = HeightOriginalDT, fill = "#482677FF", color = "#482677FF")) +
  geom_density_ridges_gradient(scale = 3, rel_min_height = 0.01,  stat="binline", bins=10) +
  # scale_fill_viridis(name = "freq", option = "C") +
  scale_x_continuous(range(0, 1)) +
  theme(legend.position = "none", text = element_text(size = 16, family = "LM Roman 10")) +
  labs(x=NULL, y="Depth of the dependency tree") 

ggsave(filename = "Figures/bloat_treedeep_ridges.pdf", plot = bloat_treedeep_ridges,
       height = 5, width = 4,  units = c("in"), device = "pdf") 

```


## Bloated dependencies in multi-module projects

```{r}
results_bloat <- results %>% 
  mutate(TypeUsage = ifelse(UsedDeclared == T, "UsedDeclared", ifelse(UsedUndeclared == T, "UsedUndeclared", ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared"))))

# is multi modulue
description_multimodule <- description %>% dplyr::select(Artifact, IsMultimodule)

# percentage of usage types for the compiled scope
filtered_results_bloat<- results_bloat %>% 
  filter(Scope %in% c("compile")) %>% 
  dplyr::group_by(Artifact, TypeUsage) %>%
  dplyr::summarise(n = n()) %>% 
  mutate (percentage = n / sum(n)) %>% 
  ungroup()

filtered_results_bloat$label <- round(filtered_results_bloat$percentage * 100, 1)

join_multimodule <- inner_join(filtered_results_bloat, description_multimodule, by = "Artifact")

means <- aggregate(percentage ~  TypeUsage, join_multimodule, mean)
means$percentage <- round(means$n, 2)
nb_direct_and_transitive_deps_boxplot <- 
  nb_direct_and_transitive_deps %>% 
  ggplot(aes(Type, n)) +
  geom_violin(trim = FALSE, aes(fill = Type)) +
  geom_boxplot(width = 0.1) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  geom_point(data = means,
             aes(x=Type, y=n),
             colour="black",
                shape=18, size=6, show.legend = FALSE)+
  geom_point(data = means,
             aes(x=Type, y=n),
             colour="lightblue", 
                shape=18, size=4, show.legend = FALSE)+
  geom_text(data = means, aes(label = n, y = n)) 

# boxplot
dodge <- position_dodge(width = 0.6)
filtered_results_bloat_multimodule_boxplot <- join_multimodule %>% 
  ggplot(aes(x = IsMultimodule, y = percentage, fill = TypeUsage)) +
   geom_violin(position = dodge) +
  geom_boxplot(width=.05, outlier.colour=NA, position = dodge, color = "black") +
    # stat_summary(fun.y=mean, geom="point", shape=23, size=2) +
    scale_fill_viridis(discrete = TRUE)  +
    ylab("%Dependencies") +
  scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) 

ggsave(filename = "Figures/filtered_results_bloat_multimodule_boxplot.pdf", plot = filtered_results_bloat_multimodule_boxplot,
       height = 4, width = 7,  units = c("in"), device = "pdf")

# count the number
join_multimodule_aggregated <- aggregate(join_multimodule$n, by=list(IsMultimodule=join_multimodule$IsMultimodule, TypeUsage=join_multimodule$TypeUsage), FUN=sum)

filtered_results_bloat_multimodule <- join_multimodule_aggregated %>% 
  ggplot(aes(x = IsMultimodule, y = x, fill = TypeUsage)) +
    geom_bar(stat="identity", position = "dodge") +
    # geom_text(size = 5, position = position_stack(vjust = 0.5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(x = "Multimodule", y = "#Dependencies") +
     scale_y_continuous(name="#Dependencies", labels = comma) +
    # scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip()

# count the percentage
join_multimodule_aggregated %>% group_by(IsMultimodule, TypeUsage) %>% summarise(x/sum(x))

join_multimodule_aggregated$IsMultimodule <- as.factor(join_multimodule_aggregated$IsMultimodule)
join_multimodule_aggregated$TypeUsage <- as.factor(join_multimodule_aggregated$TypeUsage)

join_multimodule_aggregated <- join_multimodule_aggregated %>% 
   dplyr::ungroup() %>% 
  group_by(IsMultimodule, TypeUsage) %>% 
  dplyr::mutate(freq = x / sum(x))
 
# Bar plot
filtered_results_bloat_multimodule <- join_multimodule_aggregated %>% 
  ggplot(aes(x = IsMultimodule, y = x, fill = TypeUsage)) +
    geom_bar(stat="identity", position = "dodge") +
    # geom_text(size = 5, position = position_stack(vjust = 0.5), family = "LM Roman 10") +
    scale_fill_viridis(discrete = TRUE) +
    labs(x = "Multimodule", y = "#Dependencies") +
     scale_y_continuous(name="#Dependencies", labels = comma) +
    # scale_y_continuous(labels = scales::percent) +
    theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10")) +
    coord_flip()

ggsave(filename = "Figures/filtered_results_bloat_multimodule.pdf", plot = filtered_results_bloat_multimodule,
       height = 3, width = 7,  units = c("in"), device = "pdf")
```


## Rotten bloated dependencies

```{r}
results %>% filter(AllDeps == "org.eclipse.core:filesystem:test")

View(results %>% group_by(Artifact) %>% summarise(n =n()) %>% arrange(desc(n)))

View(results %>% filter(Artifact == "org.jxls:jxls-poi:1.0.15"))

View(results %>% filter(BloatedDeclared == TRUE & Scope == "compile"))


results %>% filter(Artifact == "org.jxls:jxls-poi:1.0.15" & Scope == "compile") %>% filter(Used == F & Type == "transitive")
View(data %>% filter(GroupId == "org.apache.flink"))

results %>% group_by(Artifact) %>% summarise(n = n()) %>%  arrange(desc(n))

# org.apache.accumulo:accumulo-core:1.9.0
# org.apache.hive:hive-contrib:2.3.3
# org.apache.flink:flink-hbase_2.11:1.5.0
# org.apache.drill:drill-yarn:1.14.0
# com.netflix.dynomitemanager:dynomitemanager-common:2.0.13
# com.facebook.presto:presto-main:0.192
# ml.dmlc:xgboost4j-spark:0.72
# org.apache.hive.shims:hive-shims-0.23:2.3.3

filtered_results_bloat %>% filter(Artifact == "org.apache.accumulo:accumulo-core:1.9.0")

accumulo_versions_list <- data %>% 
  filter(GroupId == "org.apache.accumulo" & ArtifactId == "accumulo-core") %>% 
  arrange(desc(Version)) %>% 
  mutate(Artifact = paste(paste(GroupId, ArtifactId, sep = ":"), Version, sep = ":")) %>% 
  select(Artifact)

# write the list versions of the candidate library
write.csv(accumulo_versions_list, file = "/home/cesarsv/Documents/accumulo_versions_list.csv", col.names = FALSE, row.names = F, quote = F)

# read the dependency analysis results of all versions
accumulo_versions_list_results <- read_csv("Data/accumulo_versions_list_results.csv", col_names = T)

# reshape the dataset
accumulo_versions_list_results <- accumulo_versions_list_results %>%
  mutate(
    UsedDeclared =  ifelse(Used == T &
                             Declared == T, T, F),
    UsedUndeclared =  ifelse(Used == T &
                               Declared == F, T, F),
    BloatedDeclared =  ifelse(Used == F &
                                Declared == T, T, F),
    BloatedUndeclared =  ifelse(Used == F &
                                  Declared == F, T, F)
  ) %>%
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>% 
  filter(Scope == "compile" & Optional == F) %>% 
  group_by(Artifact, TypeUsage) %>% 
  summarise(n = n()) %>% 
  filter(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared")

# Extract release date for each version
data2 <- data %>% 
  filter(GroupId == "org.apache.accumulo" & ArtifactId == "accumulo-core") %>% 
  mutate(Artifact = paste(paste(GroupId, ArtifactId, sep = ":"), Version, sep = ":")) %>% 
  select(Artifact, Version, Release) 
  
accumulo_versions_list_results_plot <- inner_join(accumulo_versions_list_results, data2, by = "Artifact") %>% 
  ggplot(aes(x = Release, y = n, color = TypeUsage, shape = TypeUsage)) +
  geom_point(size = 3) +
  geom_line(aes(type = TypeUsage)) +
  geom_text(aes(label=Version), hjust=0, vjust=-0.5,show.legend = FALSE, size = 2) +
  xlab("Release Date") + 
  ylab("Count") + 
  # scale_color_viridis(discrete=TRUE) +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y",  limits = as.Date(c('2012-01-01','2019-01-01'))) +
  scale_y_continuous(limits=c(0, 100)) +
  scale_color_manual(values = c("#482173", "#2e6f8e")) +
  theme(legend.title = element_blank())

# viridis color palette: #482173, #2e6f8e, #29af7f, #bddf26
ggsave(filename = "Figures/accumulo_versions_list_results_plot.pdf", plot = accumulo_versions_list_results_plot,
       height = 3, width = 7,  units = c("in"), device = "pdf")
  
```

# To what extend dependency conflicts are superflous due to bloated dependencies?

```{r}
dependency_status <- results %>%
  filter(Scope == "compile") %>% 
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>%
  dplyr::select(Artifact, InConflict, TypeUsage) %>%
  group_by(Artifact, InConflict, TypeUsage) %>%
  summarise(n = n()) %>% 
  group_by(InConflict, TypeUsage) %>%
  summarise(sum = sum(n))

dependency_status$InConflict <-
  factor(
    dependency_status$InConflict,
    levels = c('In conflict', 'Omitted for duplicate', 'Omitted for cycle', 'Not in conflict'),
    ordered = TRUE
  )

dependency_status

dependency_status_barplot <-
  dependency_status %>% 
  filter(InConflict !=  'Omitted for cycle') %>% 
  ggplot(aes(x = InConflict, y = sum, fill=TypeUsage)) +
  geom_bar(stat="identity", position=position_dodge()) +
  geom_text(aes(label=sum), position=position_dodge(width=0.9), vjust=-0.25, size = 3.5, family = "LM Roman 10") +
  xlab(NULL) +
  ylab("#Dependencies") +
  scale_y_continuous(labels = comma) +
   coord_cartesian(ylim = c(0, 210000)) +
  scale_fill_viridis(discrete = TRUE) +
  theme(legend.position = "top", legend.title = element_blank(), text = element_text(size = 16, family = "LM Roman 10"))

ggsave(filename = "Figures/dependency_status_barplot.pdf", plot = dependency_status_barplot,
       height = 3, width = 7,  units = c("in"), device = "pdf") 

```

## Distribution of dependencies in conflict/duplicated/cycle or NO

```{r}
deps_in_conflict <- results %>%
  filter(InConflict !=  'Omitted for cycle' & Scope == "compile") %>% 
  mutate(TypeUsage = ifelse(
    UsedDeclared == T,
    "UsedDeclared",
    ifelse(
      UsedUndeclared == T,
      "UsedUndeclared",
      ifelse(BloatedDeclared == T, "BloatedDeclared", "BloatedUndeclared")
    )
  )) %>%
  group_by(Artifact, InConflict, TypeUsage) %>%
  summarise(n = n())

deps_in_conflict$InConflict <-
  factor(
    deps_in_conflict$InConflict,
    levels = c('In conflict', 'Omitted for duplicate', 'Not in conflict'),
    ordered = TRUE
  )

deps_in_conflict_debloated <- deps_in_conflict %>% filter(TypeUsage != "BloatedDeclared" & TypeUsage != "BloatedUndeclared")
 
dodge <- position_dodge(width = 0.6)
dist_deps_in_conflict_bloat_debl <- deps_in_conflict %>% 
  
  filter(InConflict == "In conflict") %>% 
  mutate(BloatStatus = ifelse(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared", "Bloated", "Debloated")) %>% 
  ggplot(aes(x = InConflict, y = n, fill = BloatStatus)) +
  geom_violin(width = 0.77, position = dodge) +
  geom_boxplot(width = .05, color = "black", position = dodge) +
  # stat_summary(fun.y=mean, geom="point", shape=23, size=2) +
  # scale_y_log10(
  #   breaks = scales::trans_breaks("log10", function(x)
  #     10 ^ x),
  #   labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  # ) +
  scale_y_continuous(breaks=c(0, 1,10,100,1000), trans="log1p") +
  ylab("#Dependencies (log)") +
  xlab(NULL) +
  scale_fill_viridis(discrete = TRUE, option = "magma", begin = 0.6, end = 0.9) +
  theme(legend.title = element_blank()) 


ggsave(filename = "Figures/dist_deps_in_conflict_bloat_debl.pdf", plot = dist_deps_in_conflict_bloat_debl,
       height = 4, width = 7,  units = c("in"), device = "pdf") 

bloated <- deps_in_conflict %>% 
  mutate(BloatStatus = ifelse(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared", "Bloated", "Debloated")) %>% 
  filter(InConflict == "In conflict" & BloatStatus == "Bloated")
summary(bloated$n)

not_bloated <- deps_in_conflict %>% 
  mutate(BloatStatus = ifelse(TypeUsage == "BloatedDeclared" | TypeUsage == "BloatedUndeclared", "Bloated", "Debloated")) %>% 
  filter(InConflict == "In conflict" & BloatStatus == "Debloated")
summary(not_bloated$n)
```


## Comparison between the height of the original vs debloated dependency tree

```{r}
description$HeightOriginalDT <-  as.numeric(description$HeightOriginalDT)
description$HeightDebloatedDT <-  as.numeric(description$HeightDebloatedDT)

description %>% ggplot(aes(HeightOriginalDT)) +
  geom_histogram()

description %>% ggplot(aes(HeightDebloatedDT)) +
  geom_histogram()

metrics %>%
  filter(is.na(isPassive) == F) %>%
  mutate(Type = ifelse(isPassive == "true", "passive", "active")) %>% 
  # filter(Type == "passive") %>% 
  # filter(validityPeriod > 0) %>% 
  ggplot(aes(Type, lifespan)) +
  geom_violin(trim = FALSE, fill = "#CCCCCC") +
  geom_boxplot(width = 0.1) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x)
      10 ^ x),
    labels = scales::trans_format("log10", scales::math_format(10 ^ .x))
  ) +
  coord_flip() +
  theme(legend.position = "none") +
  xlab(NULL) +
  ylab(TeX("lifespan (log_{10} scale)"))
  
```


```{r}
# library
library(ggridges)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)

# Load dataset from github
data <- read.table("https://raw.githubusercontent.com/zonination/perceptions/master/probly.csv", header=TRUE, sep=",")
data <- data %>% 
  gather(key="text", value="value") %>%
  mutate(text = gsub("\\.", " ",text)) %>%
  mutate(value = round(as.numeric(value),0)) %>%
  filter(text %in% c("Almost Certainly","Very Good Chance","We Believe","Likely","About Even", "Little Chance", "Chances Are Slight", "Almost No Chance"))

# Plot
data %>%
  mutate(text = fct_reorder(text, value)) %>%
  ggplot( aes(y=text, x=value,  fill=text)) +
    geom_density_ridges(alpha=0.6, stat="binline", bins=20) +
    theme_ridges() +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
    xlab("") +
    ylab("Assigned Probability (%)")
```


```{r}
# data preparation library(ggplot2)
col <- c("red", "pink", "orange", "brown")
freq <- c(101, 68, 34, 18)

tmp %>% group_by(Artifact,Type) %>% 
  summarise(count=n())
## create data frame
colour.df <- data.frame(col, freq)
colour.df

## calculate percentage 
colour.df$percentage = colour.df$freq / sum(colour.df$freq)* 100
colour.df = colour.df[rev(order(colour.df$percentage)), ]
colour.df$ymax = cumsum(colour.df$percentage)
colour.df$ymin = c(0, head(colour.df$ymax, n = -1))
colour.df

library(ggrepel)
colour.df %>% ggplot(aes(fill = col, ymax = ymax, ymin = ymin, xmax = 100, xmin = 80)) +
    geom_rect(colour = "black") +
    coord_polar(theta = "y") + 
    xlim(c(0, 100)) +
    geom_label_repel(aes(label = paste(round(percentage,2),"%"), x = 100, y = (ymin + ymax)/2),inherit.aes = F, show.legend = F, size = 5)+
    theme(legend.title = element_text(colour = "black", size = 16, face = "bold"), 
        legend.text = element_text(colour = "black", size = 15), 
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()) +
    annotate("text", x = 0, y = 0, size = 8, label = "Dependencies") +
  theme(legend.position = "top",
        text = element_text(size = 16, family = "LM Roman 10"))
    
```

## Plot conflicts

```{r}
results <-
  results %>% 
  filter(Type == "transitive") %>% 
  mutate(UsageType = ifelse(
    UsedDeclared == "true",
    "UsedDeclared",
    ifelse(
      UsedUndeclared == "true",
      "UsedUndeclared",
      ifelse(BloatedDeclared == "true", "BloatedDeclared", "BloatedUndeclared")
    )
  )) 

results %>% filter(UsageType == "UsedUndeclared" | UsageType == "BloatedUndeclared") %>% 
  ggplot(aes(UsageType, ..count..)) + geom_bar(aes(fill = InConflict), position = "dodge") +
   theme(legend.position = "top", text=element_text(size=14, family="LM Roman 10"))

ggsave(filename = "Figures/usage-conflicts.pdf", height = 4, width = 6,  units = c("in"), device = cairo_pdf)

```


# How much can developers benefit from removing their direct bloated dependencies?

```{r}
results %>% 
  select(Artifact, AllDeps, )
```

